\documentclass[11pt]{article}

% --- packages ---
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{float}
\usepackage[numbers]{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
% \usepackage{microtype}  % disabled: MiKTeX font expansion issue
\usepackage{bookmark}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black
}

% --- metadata ---
\title{%
  The Learnable Bridge:\\
  Task Fingerprinting and Adapter Composition\\
  via Structured Coupling in Low-Rank Adaptation%
}

\author{%
  Timothy Paul Bielec%
  \thanks{Corresponding author. \texttt{timothy@promptcrafted.com}}%
  \\
  Promptcrafted LLC%
}

\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
We insert a learnable $6 \times 6$ coupling matrix---the \emph{bridge}---between
the down- and up-projections of a LoRA adapter, adding 36 parameters per
adapted layer ($0.02\%$ overhead). The bridge begins as the identity matrix,
recovering standard LoRA exactly. Through seven experiments on Qwen2.5
(1.5B and 7B), we establish four findings: (1)~bridge matrices encode task
identity---a leave-one-out SVM on 28 query-projection bridges (1,008
parameters) classifies task type at 82.1\% accuracy (rising to 84.5\% at
the final checkpoint) versus 33.3\% chance; (2)~FCC lattice topology learns
$4.6\times$ more cross-channel coupling than cubic topology, though neither
improves training loss; (3)~linear interpolation of bridge matrices between
adapters preserves eigenspectrum structure (cosine $> 0.999$), enabling
low-dimensional adapter composition analysis; (4)~contrastive bridge
pre-training installs directional preference (co-planar/cross-planar ratio
$> 2{,}600\times$) that persists as a stable geometric memory after
supervision is removed---decaying to $16\times$ over 6{,}000 unsupervised
steps but never reaching the null baseline of $1.0$, while overall bridge
connectivity rises $3.7\times$ independently. The bridge does not improve
loss or perplexity relative to standard LoRA at matched rank. Its
contribution is diagnostic, not generative: a 36-parameter summary of
adapter coupling structure, readable without inference or evaluation.
\end{abstract}

% --- Sections ---
\input{sections/section1_introduction}
\input{sections/section2_architecture}

% Section 3: Theoretical Grounding
\section{Theoretical Grounding}
\label{sec:theory}

The bridge's diagnostic utility rests on a simple observation about
gradient flow. The loss gradient with respect to bridge element $M_{ij}$
is proportional to the co-activation between down-projection channel $j$
and up-projection gradient channel $i$ (Eq.~\ref{eq:bridge-grad}).
Bridge entries therefore accumulate a running summary of inter-channel
correlation as shaped by the training objective.

Three properties of the bridge gradient make it diagnostic:

\paragraph{Selectivity.}
If training data induces systematic correlations between specific channel
pairs---because certain feature combinations co-occur across examples---the
corresponding off-diagonal entries grow. If channels are independent,
$M$ remains near $I_C$. The bridge is shaped by the task's statistical
structure, not by initialization or architectural bias.

\paragraph{Compactness.}
The bridge occupies $C^2 = 36$ parameters per layer, regardless of
model dimension $d_\text{model}$. This is smaller than a single hidden
state vector in the base model. The diagnostic compression ratio is
$d_\text{model}^2 / C^2 \approx 3.6 \times 10^5$ for Qwen2.5-7B.

\paragraph{Invariant diagnostics.}
While bridge entries depend on the optimizer's choice of rank-space
rotation, the eigenspectrum of $M$ is rotation-invariant. Two bridges
trained on the same task with different random seeds will differ
entry-wise but converge to similar eigenspectra (cosine similarity
$> 0.999$ across all task pairs in our experiments). Spectral analysis
of the bridge is therefore reproducible across training runs.

The theoretical prediction---that task structure flows into bridge
structure through gradient co-activation---is confirmed empirically
in Sections~\ref{sec:exp-fingerprint} and~\ref{sec:exp-merging}.

\input{sections/section4_experiments}

% Section 5: Related Work
\section{Related Work}
\label{sec:related}

\paragraph{LoRA variants.}
LoRA~\cite{hu2022lora} established the $B \cdot A$ factorization for
parameter-efficient fine-tuning. Subsequent work has modified this
factorization in several directions. AdaLoRA~\cite{zhang2023adalora}
dynamically adjusts rank allocation across layers via importance
scoring. LoRA-XS~\cite{banaei2024loraxs} inserts an $r \times r$
matrix between frozen $A$ and $B$ projections, achieving extreme
parameter reduction by learning only the intermediate coupling.
MELoRA~\cite{ren2024melora} partitions rank dimensions into
independent mini-LoRAs, equivalent to a block-diagonal bridge with
frozen identity blocks. DoRA~\cite{liu2024dora} decomposes the
update into magnitude and direction components. None of these variants
use the intermediate representation as a \emph{diagnostic} of adapter
behavior; all frame intermediate structure as a means to improve
adaptation efficiency.

\paragraph{Adapter merging.}
Task arithmetic~\cite{ilharco2023editing} composes adapters by
linear combination of weight deltas. TIES-Merging~\cite{yadav2023ties}
prunes small-magnitude parameters and resolves sign conflicts before
averaging. DARE~\cite{yu2024dare} randomly drops adapter parameters
before merging. AdaMerging~\cite{yang2024adamerging} learns per-layer
merging coefficients. These methods operate on full weight matrices
or their SVD factors without a compact representation of task identity.
Bridge interpolation (\S\ref{sec:exp-merging}) provides such a
representation: a 36-dimensional space in which adapter compatibility
can be assessed before committing to an expensive full-weight merge.

\paragraph{Adapter analysis.}
Prior work on understanding LoRA representations has focused on
weight-space structure. The LoRA training dynamics
analysis~\cite{biderman2024lora} examines how $A$ and $B$ evolve
during training but provides no compact summary of the learned
coupling. Spectral analysis of adapter weights~\cite{zhang2023adalora}
operates on the full $d_\text{out} \times r$ projection matrices.
Our bridge-based diagnostics compress the coupling structure to
$C^2 = 36$ parameters per layer while retaining sufficient information
for task classification at 84.5\% accuracy.

\input{sections/section6_discussion}

\section*{Acknowledgments}
This work was conducted entirely on consumer and cloud GPU hardware:
NVIDIA RTX 6000 Ada 48\,GB (local), RTX 4090 16\,GB and RTX 3090
24\,GB (RunPod) for small-scale and contrastive experiments. The \texttt{rhombic} library and all experimental code are
available at \url{https://github.com/promptcrafted/rhombic} under
MPL-2.0. Bridge matrices for all experiments are included as NumPy
arrays for full reproducibility.

\bibliographystyle{plainnat}
\bibliography{refs-paper3}

\end{document}
